Can Exposure To Forgiving AI Foster Cooperative Play?

### 2024-03-27 article.Rmd
v2 article

In this study, we use a randomized control trial to test a manipulation aimed at repairing potential breakdowns in RTG cooperation. Participants are exposed to agents designed with a limited propensity for retaliation, potentially mitigating ingrained mistrust. We stratify our sample based on rejection sensitivity (RS): a tendency to anxiously expect, readily perceive, and intensely react to rejection. It has been identified as a potential mechanism linking early interpersonal trauma to negative mental health outcomes [@downey_early_1997]. Extensive research demonstrates strong associations between rejection sensitivity and various mental health conditions, including depression, anxiety, personality disorders, and self-harm (for a review, see @gao_associations_2017). However little is known about whether rejection sensitivity is linked to mistrust and difficulties in cooperation in social dilemmas. Finally, the computerized investor in the RTG was programmed to play according to a hidden Markov model estimated from real players' data in prior research. A key aspect of this model is that the actions of the investor depend on a latent "trust state" which reacts dynamically to the trustees' returns simulating real-life trust-building scenarios. Such a model, informed by empirical data and integrating a responsive trust mechanism, represents a novel approach to study interactive behavior in multi-player games whilst keeping a high degree of experimental control.

@@

@@

A total of 206 participants were recruited on the Prolific Academic platform (prolific.co).
<!-- The mean age of participants was `r round(mean(as.numeric(d_finished$exitSurvey.age), na.rm=TRUE),1)` years, with a `r round(sd(as.numeric(d_finished$exitSurvey.age), na.rm=TRUE),1)` years standard deviation.  -->
Participants were paid a fixed fee of £6 plus a bonus payment dependent on their performance that averaged £XXX. Participants were pre-screened on Prolific using the Rejection Sensitivity Questionnaire to form two similarly sized groups: One with high (RSQ score > 15) and the other with low rejection sensitivity (RSQ score < 10).

@@

The model was estimated using the `afex` package [@singmann_afex_2022] in R. More complex models with additional random effects could not be estimated reliably, and as such the estimated model can be considered to include the optimal random effects structure [@matuschek_balancing_2017]. A similar process was used to establish the random effects structures of other linear mixed-effects models used throughout the statistical analyses. For the $F$-tests, we used the Kenward-Roger approximation to the degrees of freedom, as implemented in the R package "afex". We Z-transform the Investment variable as centering is beneficial to interpreting the main effects more easily in the presence of interactions.

@@

# Pivot longer ratings on each attribute 
df_coop <- d_finished %>% dplyr::select(playerId,condition.f,contains("cooperative")) %>%  
  pivot_longer(cols=contains("cooperative"), names_to = c("game"), names_pattern ="rating_cooperative_(.*)", values_to = "rating_coop") %>% distinct()

df_forgiv <- d_finished %>% dplyr::select(playerId,condition.f,contains("forgiving")) %>%
  pivot_longer(cols=contains("forgiving"), names_to = c("game"), names_pattern ="rating_forgiving_(.*)", values_to = "rating_forgiving") %>% distinct()

df_playAgain <- d_finished %>% dplyr::select(playerId,condition.f,contains("again")) %>%  
  pivot_longer(cols=contains("again"), names_to = c("game"), names_pattern ="rating_playAgain_(.*)", values_to = "rating_playAgain") %>% distinct()

#merge all data frames together
datRatings <- list(df_coop, df_forgiv,df_playAgain) %>% 
              reduce(full_join, by=c('playerId','game','condition.f')) %>%
              mutate(gameNum.f = factor(game,labels = c("pre", "expo1", "expo2", "expo3","post"),
                                      levels=c("1", "11", "12", "13", "2")))

@@

temp2 <- d_finished %>% dplyr::filter(gameNum.f %in% c("pre", "post")) %>%
  dplyr::select(id, payoffTrust1, payoffTrust2, condition.f) %>%
  unique()

@@

cor(temp2$payoffTrust1,temp2$payoffTrust2)

@@

@@

df_long_ratings <- datRatings %>%
  pivot_longer(
    cols = starts_with("rating_"), 
    names_to = "rating_type", 
    values_to = "rating_value"
  ) %>%
  dplyr::select(playerId, condition.f, gameNum.f, rating_type, rating_value)

# First, calculate the mean and SEM for each group
summary_df <- df_long_ratings %>%
  group_by(gameNum.f, condition.f, rating_type) %>%
  summarise(
    mean = mean(rating_value, na.rm = TRUE),
    sem = sd(rating_value, na.rm = TRUE)/sqrt(n())
  ) %>%
  ungroup()

@@

avg_ret_df <- d_finished %>%
  dplyr::select(playerId,roundType,investment,returns,roundNum,gameNum.f,condition.f, phase.f, high_RS) %>% 
  filter(roundType=="trust",!is.na(roundNum)) %>% 
  mutate(roundNum = as.numeric(as.character(roundNum))) %>%
  mutate(inv_scaled = as.vector(scale(investment))) %>% 
  mutate(inv_pct = investment/20, ret_pct_0 = ifelse(investment==0,0,returns/(3*investment)),ret_pct_na = ifelse(investment==0,NA,returns/(3*investment)))

@@

mod_rating_coop <- mixed( rating_coop ~ gameNum.f*condition.f + (1 | playerId), datRatings, REML= TRUE, method="KR")

#summary(mod_rating_coop)

@@

@@

mod_rating_forg <- mixed( rating_forgiving ~ gameNum.f*condition.f + (1 | playerId), datRatings, REML= TRUE, method="KR")

@@

@@

mod_rating_again <- mixed( rating_playAgain ~ gameNum.f*condition.f + (1 | playerId), datRatings, REML= TRUE, method="KR")

@@

@@

Figure \@ref(fig:plotRatings) shows participants' ratings of each player they faced by condition. Compared to the first RTG game, participants in the manipulation condition rated the investors they faced in the exposure phase (the forgiving AI) as more Cooperative `r papaja::apa_print(comp2_coop)$full_result$Manipulation_PrevsExpo`. No difference in ratings on forgiveness and whether they would like to face them again. Those in the control condition rated the investors faced in the exposure group (same HMM) as less cooperative `r papaja::apa_print(comp2_coop)$full_result$Control_PrevsExpo`, less forgiving `r papaja::apa_print(comp2_forg)$full_result$Control_PrevsExpo`, and were less keen on facing them again `r papaja::apa_print(comp2_again)$full_result$Control_PrevsExpo`.

```{r plotRatings,include=T, echo=FALSE, fig.cap="Averages and standard errors of the participants ratings of the opponent for each game and condition. Pre and post are the 15 round repeated trust games before and after the exposure phase respectively. The games titled expo 1 to 3 are the three 7 round games during the exposure phase. We note that absent the exposure to the forgiving AI, the ratings get worse on aggregate even through the participant faces the same human like HMM. In the manipulation condition, participants rate the nicer HMM as more cooperative but not more forgiving. When they face the human-like HMM again, its rating are considerably worse. ",fig.align="center", fig.width=6, fig.height = 6}
# Now create a plot using geom_errorbar for SEM and geom_point for the mean
ggplot(summary_df, aes(x = gameNum.f, y = mean, group = rating_type, color = rating_type)) +
  geom_errorbar(aes(ymin = mean-sem, ymax = mean+sem), width = 0.2) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  facet_wrap(~~rating_type*condition.f, nrow = 3, ncol = 2, scales = "free") +
  labs(x = "Game Number", y = "Mean Rating") +
  theme_minimal() +
  theme(legend.position = "bottom")

@@

Comparing post-exposure to exposure ratings, participants in the manipulation condition rated the investors they faced in the exposure phase as less cooperative`r papaja::apa_print(comp3_coop)$full_result$Manipulation_PostvsExpo`, less forgiving `r papaja::apa_print(comp3_forg)$full_result$Manipulation_PostvsExpo`, and were less willing to face them again `r papaja::apa_print(comp3_again)$full_result$Manipulation_PostvsExpo`. Ratings for those in the control group did not differ on how cooperative, forgiving the investors were, and on willingness to face them again.

Finally, comparing ratings of the investors before and after the exposure phase, those in the manipulation condition rated them similarly on cooperation, lower on forgiveness `r papaja::apa_print(comp1_forg)$full_result$Manipulation_PrevsPost` and lower on willingness to face them again `r papaja::apa_print(comp1_again)$full_result$Manipulation_PrevsPost`. Those in the control condition rated the investors post the exposure phase lower on all three attributes.

@@

@@

# 'inv_scaled' is the continuous predictor, and 'outcome' is your dependent variable.
ggplot(prepost_df, aes(x = inv_scaled, y = ret_pct_na, color = phase.f)) +
  geom_point(alpha = 0.6) + # Adjust point transparency with alpha
  geom_smooth(method = "lm", se = FALSE) + # Add linear regression lines without confidence intervals
  labs(title = "Effect of inv_scaled on Outcome by Condition",
       x = "inv_scaled", y = "Outcome") +
  theme_minimal() +
  facet_wrap(~condition.f)

@@

@@

Mixed-effects analysis on the percentage returns shows a significant main effect of Phase (Pre vs. Post RTG game), `r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f`, with higher percentage returns in the first RTG compared to the second. Importantly, we also find an interaction between Condition and Phase (RTG pre- vs. post-intervention), `r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f`. Post-hoc tests show a decrease in the percentage returned only in the intervention condition, pre - post,
`r papaja::apa_print(prepost_bycond)$full_result$Manipulation_Pre_post`, but no change in the control condition. Looking at this interaction effect for the two levels of RS, we find a three way interaction between Phase, Condition and RS, such that this decrease in returns in the manipulation condition is present in the low RS group `r papaja::apa_print(contrasts)$full_result$Manipulation_LowRS_Pre_post` but not in the high RS group.

```{r gamesPlot, include=T, echo=FALSE, fig.cap="Averages and standard errors of the trustee's return as a percentage of the multiplied investment received by Condition, Phase, and game round. The blue line shows the returns pre-manipulation and the green line post-manipulation. We note a different reaction to the pre-programmed one-off low investment between the two conditions: Whilst there is a dip in returns pre-manipulation for both conditions,  post manipulation we see higher returns in the manipulation condition compared to the dip in returns seen in the control condition in the right panel",fig.align="center", fig.width=6, fig.height = 6}

@@

# Plot
ggplot(avg_ret_df %>% filter(gameNum.f %in% c("pre","post")), aes(x=as.factor(roundNum), y=ret_pct_na, group=gameNum.f, color = gameNum.f, fill=gameNum.f)) +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "ribbon", aes(ymin=..ymin.., ymax=..ymax..),
               alpha = 0.3, linetype = 0) +
  geom_vline(data = vline_data, aes(xintercept = xintercept, linetype = Defection_round), color = "black", linewidth = 0.6) +
  scale_linetype_manual(values = c("Pre-manipulation" = "dotted", "Post-manipulation" = "dashed"),name = "Defection round") +
  labs(x = "Round",
       y = "Percentage Return",
       color = "Trust Game Phase") +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.box = "vertical", # This will stack the legends vertically
        legend.key = element_blank(),
        legend.margin = margin(t = 0.2, b = 0.2, unit = "pt"),
        legend.box.margin = margin(t = 0, b = 0, unit = "pt"),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10)) +
  scale_color_manual(values = c("darkblue", "darkgreen"),
                     labels = c("Pre-manipulation", "Post-manipulation")) +
  scale_fill_manual(values = c("darkblue", "darkgreen"),
                    guide = "none") + # Hide the fill legend
  guides(color = guide_legend(override.aes = list(fill = NA)))+  # Remove fill for color legend keys
  guides(color = guide_legend(title = "Participant Returns Proportions", override.aes = list(fill = NA)),
         linetype = guide_legend(title = "Defection round")) +
  facet_wrap(~high_RS*condition.f)

@@

There was also a significant main effect of Investment, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$inv_scaled)`,
such that higher investments were associated with higher percentage returns indicating positive reciprocity. An Investment by Condition interaction, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)`, indicates that returns were more affected by investments in the control condition. We also find a three way interaction between Phase,  Investment and RS
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$gameNum_f_condition_f_inv_scaled)`, showing that the differentiated effect of the investment on the proportion returned by RS group is itself moderated by the Phase (pre- vs post intervention).

>>>>

In this study, we use a randomized control trial to test an exposure manipulation aimed at repairing potential breakdowns in RTG cooperation. We stratify the sample of participants based on rejection sensitivity (RS): a tendency to anxiously expect, readily perceive, and intensely react to rejection. It has been identified as a potential mechanism linking early interpersonal trauma to negative mental health outcomes [@downey_early_1997]. Extensive research demonstrates strong associations between rejection sensitivity and various mental health conditions, including depression, anxiety, personality disorders, and self-harm (for a review, see @gao_associations_2017). However little is known about whether rejection sensitivity is linked to mistrust and difficulties in cooperation in social dilemmas.

Participants are exposed to artificial agents designed with a limited propensity for retaliation, potentially mitigating ingrained mistrust. The computerized investor in the RTG was programmed to play using a generative model of how humans play fit to real players' data from prior studies. A key aspect of this model is that the actions of the investor depend on a latent "trust state" which reacts dynamically to the trustees' returns simulating real-life trust-building scenarios. An advantage of having a generative model of behavior is the possibility of controlling different aspects of the agent's strategy such as its general policy, the propensity to cooperate actively or the propensity to retaliate after breakdowns of cooperation. Such generative  model of behavior, informed by empirical data and integrating a responsive trust mechanism, represent a novel approach to study interactive behavior in multi-player games whilst keeping a high degree of experimental control.

@@

save(d_finished, file = "data/d_finished.RData")

@@

A total of 206 participants were recruited on the Prolific Academic platform (prolific.co).
<!-- The mean age of participants was `r round(mean(as.numeric(d_finished$exitSurvey.age), na.rm=TRUE),1)` years, with a `r round(sd(as.numeric(d_finished$exitSurvey.age), na.rm=TRUE),1)` years standard deviation.  -->
Participants were paid a fixed fee of £6 plus a bonus payment dependent on their performance that averaged £0.5. Participants were pre-screened on Prolific using the Rejection Sensitivity Questionnaire to form two similarly sized groups: One with high (RSQ score > 15) and the other with low rejection sensitivity (RSQ score < 10).

@@

The model was estimated using the `afex` package [@singmann_afex_2022] in R. More complex models with additional random effects could not be estimated reliably, and as such the estimated model can be considered to include the optimal random effects structure [@matuschek_balancing_2017]. A similar process was used to establish the random effects structures of other linear mixed-effects models used throughout the statistical analyses. For the $F$-tests, we used the Kenward-Roger approximation to the degrees of freedom, as implemented in the R package "afex". We Z-transform the Investment variable (subtract the overall investment mean and divide by overall standard deviation) as centering is beneficial to interpreting the main effects more easily in the presence of interactions.

@@

# Pivot longer ratings on each attribute 
df_coop <- d_finished %>% dplyr::select(playerId,condition.f,contains("cooperative"),high_RS) %>%  
  pivot_longer(cols=contains("cooperative"), names_to = c("game"), names_pattern ="rating_cooperative_(.*)", values_to = "rating_coop") %>% distinct()

df_forgiv <- d_finished %>% dplyr::select(playerId,condition.f,contains("forgiving"),high_RS) %>%
  pivot_longer(cols=contains("forgiving"), names_to = c("game"), names_pattern ="rating_forgiving_(.*)", values_to = "rating_forgiving") %>% distinct()

df_playAgain <- d_finished %>% dplyr::select(playerId,condition.f,contains("again"),high_RS) %>%  
  pivot_longer(cols=contains("again"), names_to = c("game"), names_pattern ="rating_playAgain_(.*)", values_to = "rating_playAgain") %>% distinct()

#merge all data frames together
datRatings <- list(df_coop, df_forgiv,df_playAgain) %>% 
              reduce(full_join, by=c('playerId','game','condition.f','high_RS')) %>%
              mutate(gameNum.f = factor(game,labels = c("pre", "expo1", "expo2", "expo3","post"),
                                      levels=c("1", "11", "12", "13", "2")))

@@

temp2 <- d_finished %>% dplyr::filter(gameNum.f %in% c("pre", "post")) %>%
  dplyr::select(id, payoffTrust1, payoffTrust2, condition.f, RS_score, LPFS_score) %>%
  mutate(perf = payoffTrust1 + payoffTrust2) %>% 
  unique()

# #                            Bonus Calculation
# avg_perf <- mean(temp2$perf)
# lowRS_bonus <- temp2 %>% filter(high_RS== "low RS") %>% mutate(bonus= round(0.5*perf/avg_perf,2)) %>% dplyr::select(id, bonus)
# write.csv(lowRS_bonus,"data/bonusLow.csv")
# 
# highRS_bonus <- temp2 %>% filter(high_RS== "high RS") %>% mutate(bonus= round(0.5*perf/avg_perf,2)) %>% dplyr::select(id, bonus)
# write.csv(highRS_bonus,"data/bonusHigh.csv")

@@

@@

```{r}
# Sanity checks

# prolific_highRS <- read.csv(file="/Users/mbp17/Downloads/prolific_export_highRS.csv")
# 
# # Which Ids in highRS file but not in Prolific? 
# 
# # Find IDs in highRS_bonus that are not in prolific_highRS
# ids_in_highRS_not_in_prolific <- setdiff(highRS_bonus$id, prolific_highRS$Participant.id)
# 
# # Find IDs in prolific_highRS that are not in highRS_bonus
# ids_in_prolific_not_in_highRS <- setdiff(prolific_highRS$Participant.id, highRS_bonus$id)
# 
# # ids_in_highRS_not_in_prolific now contains IDs that are in highRS_bonus but not in prolific_highRS
# # ids_in_prolific_not_in_highRS now contains IDs that are in prolific_highRS but not in highRS_bonus
# 
# # print these variables to see the IDs
# print(ids_in_highRS_not_in_prolific)
# print(ids_in_prolific_not_in_highRS)
```

@@

df_long_ratings <- datRatings %>%
  pivot_longer(
    cols = starts_with("rating_"), 
    names_to = "rating_type", 
    values_to = "rating_value"
  ) %>%
  dplyr::select(playerId, condition.f, gameNum.f, rating_type, rating_value, high_RS)

# First, calculate the mean and SEM for each group
summary_df <- df_long_ratings %>%
  group_by(gameNum.f, condition.f, rating_type, high_RS) %>%
  summarise(
    mean = mean(rating_value, na.rm = TRUE),
    sem = sd(rating_value, na.rm = TRUE)/sqrt(n())
  ) %>%
  ungroup()

@@

avg_ret_df <- d_finished %>%
  dplyr::select(playerId,roundType,investment,returns,roundNum,gameNum.f,condition.f, phase.f, high_RS, LPFS_score, RS_score) %>% 
  filter(roundType=="trust",!is.na(roundNum)) %>% 
  mutate(roundNum = as.numeric(as.character(roundNum))) %>%
  mutate(inv_scaled = as.vector(scale(investment))) %>% 
  mutate(inv_pct = investment/20, ret_pct_0 = ifelse(investment==0,0,returns/(3*investment)),ret_pct_na = ifelse(investment==0,NA,returns/(3*investment)))

@@

mod_rating_coop <- mixed( rating_coop ~ gameNum.f*condition.f*high_RS + (1 | playerId), datRatings, REML= TRUE, method="KR")

summary(mod_rating_coop)

@@

```{r}
# Compute the estimated marginal means (EMMs) for the interaction
emm_coop_interaction <- emmeans(mod_rating_coop, ~ gameNum.f | condition.f * high_RS)

comp1_coop_interaction <- contrast(emm_coop_interaction, list("Pre vs Post" = contrast_matrix[1,]))
comp2_coop_interaction <- contrast(emm_coop_interaction, list("Pre vs Expo" = contrast_matrix[2,]))
# comp3_coop_interaction <- contrast(emm_coop_interaction, list("Post vs Expo" = contrast_matrix[3,]))

# Output the results
print("Pre vs Post")
comp1_coop_interaction
print("Pre vs Expo")
test(comp2_coop_interaction, adjust="bonferroni")
test(comp2_coop_interaction)

# test_comp3_coop_interaction
```

@@

mod_rating_forg <- mixed( rating_forgiving ~ gameNum.f*condition.f*high_RS + (1 | playerId), datRatings, REML= TRUE, method="KR")

@@

```{r}
# Compute the estimated marginal means (EMMs) for the interaction
emm_forg_interaction <- emmeans(mod_rating_forg, ~ gameNum.f | condition.f * high_RS)

comp1_forg_interaction <- contrast(emm_forg_interaction, list("Pre vs Post" = contrast_matrix[1,]))
comp2_forg_interaction <- contrast(emm_forg_interaction, list("Pre vs Expo" = contrast_matrix[2,]))
# comp3_forg_interaction <- contrast(emm_forg_interaction, list("Post vs Expo" = contrast_matrix[3,]))

# Output the results
print("Pre vs Post")
comp1_forg_interaction
print("Pre vs Expo")
comp2_forg_interaction

# test_comp3_forg_interaction
```

@@

mod_rating_again <- mixed( rating_playAgain ~ gameNum.f*condition.f*high_RS + (1 | playerId), datRatings, REML= TRUE, method="KR")

@@

```{r}
# Compute the estimated marginal means (EMMs) for the interaction
emm_again_interaction <- emmeans(mod_rating_again, ~ gameNum.f | condition.f * high_RS)

comp1_again_interaction <- contrast(emm_again_interaction, list("Pre vs Post" = contrast_matrix[1,]))
comp2_again_interaction <- contrast(emm_again_interaction, list("Pre vs Expo" = contrast_matrix[2,]))
# comp3_again_interaction <- contrast(emm_again_interaction, list("Post vs Expo" = contrast_matrix[3,]))

# Output the results
print("Pre vs Post")
comp1_again_interaction
print("Pre vs Expo")
comp2_again_interaction

# test_comp3_again_interaction
```

@@

Figure \@ref(fig:plotRatings) shows participants' ratings of each player they faced by condition and RS group. We will focus on two contrasts to analyse the ratings, by Condition and RS group. The first is between the rating in the first phase ("pre") when participants phase the human-like HMM, and the average rating during the exposure phase (average of "expo1", "expo2" and "expo3") where they either face the forgiving HMM (Manipulation condition) or the human-like HMM again (Control condition). The second contrast is between the "pre" and "post" phases of the experiment where in both conditions participants face the same human-like HMM.

### Comparing pre and exposure ratings

For those with high Rejection Sensitivity, participants in the Manipulation condition rated the investors they faced in the exposure phase (the forgiving HMM) as more Cooperative `r papaja::apa_print(comp2_coop_interaction)$full_result$Manipulation_HighRS_PrevsExpo`. There was no difference in ratings on forgiveness and whether they would like to face the co-players again. 
Those in the control condition rated the investors faced in the exposure group (same HMM) as less cooperative `r papaja::apa_print(comp2_coop_interaction)$full_result$Control_HighRS_PrevsExpo`, less forgiving `r papaja::apa_print(comp2_forg_interaction)$full_result$Control_HighRS_PrevsExpo`, and were less keen on facing them again `r papaja::apa_print(comp2_again_interaction)$full_result$Control_HighRS_PrevsExpo`.

For those with low Rejection Sensitivity, there was no difference in any of the ratings between the "pre" and "exposure" phases for the Manipulation condition. For the Control condition, participants indicated less willingness to face the exposure co-player compared to the "pre" player, `r papaja::apa_print(comp2_again_interaction)$full_result$Control_LowRS_PrevsExpo` but also did not differ on cooperation and forgiveness ratings. In summary, those with low rejection sensitivity had a mostly undifferentiated perception of players between the pre and exposure phases, even when the co-player was in fact designed to be more forgiving. In contrast, we see that exposing participants with high Rejection Sensitivity to a more cooperative and more forgiving agent has compensated for the decrease in ratings that would have occurred if faced with a co-player with the exact same strategy.

```{r plotRatings,include=T, echo=FALSE, fig.cap="Averages and standard errors of the participants ratings of the opponent for each game and condition. Pre and post are the 15 round repeated trust games before and after the exposure phase respectively. The games titled expo 1 to 3 are the three 7 round games during the exposure phase. We note that absent the exposure to the forgiving AI, the ratings get worse on aggregate even through the participant faces the same human like HMM. In the manipulation condition, participants rate the nicer HMM as more cooperative but not more forgiving. When they face the human-like HMM again, its rating are considerably worse. ",fig.align="center", fig.width=6, fig.height = 6}

# ggplot(summary_df, aes(x = gameNum.f, y = mean, group = rating_type, color = rating_type)) +
#   geom_errorbar(aes(ymin = mean-sem, ymax = mean+sem), width = 0.2) +
#   geom_point(position = position_dodge(width = 0.5), size = 3) +
#   facet_wrap(~~rating_type*condition.f, nrow = 3, ncol = 2, scales = "free") +
#   labs(x = "Game Number", y = "Mean Rating") +
#   theme_minimal() +
#   theme(legend.position = "bottom")

ggplot(summary_df, aes(x = gameNum.f, y = mean, group = interaction(rating_type, condition.f), color = rating_type)) +
  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), width = .1, position = position_dodge(width = 0.25)) +
  geom_line(linetype = "dashed", position = position_dodge(width = 0.25)) +
  geom_point(position = position_dodge(width = 0.25), size = 3) +
  facet_wrap(~high_RS*condition.f, scales = "free_x") +
  labs(x = "Phase", y = "Mean Rating", color = "Rating Type") +
  scale_color_manual(values = c("rating_coop" = "blue", "rating_forgiving" = "red", "rating_playAgain" = "green"),
                     labels = c("Cooperative", "Forgiving", "Play again")) +
  theme_minimal() +
  theme(legend.position = "bottom")

@@

ggplot(summary_df, aes(x = gameNum.f, y = mean, group = interaction(rating_type, condition.f))) +
  geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem, color = condition.f), 
                width = .1, position = position_dodge(width = 0.25)) +
  geom_line(aes(color = condition.f), linetype = "dashed", position = position_dodge(width = 0.25)) +
  geom_point(aes(color = condition.f), position = position_dodge(width = 0.25), size = 3) +
  facet_grid(high_RS ~ rating_type, scales = "free_x") +
  scale_color_manual(values = c("Control" = "blue", "Manipulation" = "red")) +
  labs(x = "Phase", y = "Mean Rating", color = "Condition") +
  theme_minimal() +
  theme(legend.position = "bottom")

### Comparing pre and post ratings
<!-- Comparing post-exposure to exposure ratings, participants in the manipulation condition rated the investors they faced in the exposure phase as less cooperative,  -->
<!-- `r papaja::apa_print(comp3_coop)$full_result$Manipulation_PostvsExpo`, less forgiving `r papaja::apa_print(comp3_forg)$full_result$Manipulation_PostvsExpo`, and were less willing to face them again `r papaja::apa_print(comp3_again)$full_result$Manipulation_PostvsExpo`. Ratings for those in the control group did not differ on how cooperative, forgiving the investors were, and on willingness to face them again.  -->

Participants high on Rejection Sensitivity in the Manipulation condition rated the co-players in the "post" phase similarly on cooperation, lower on 
forgiveness `r papaja::apa_print(comp1_forg_interaction)$full_result$Manipulation_HighRS_PrevsPost`,
and lower on willingness to face them again `r papaja::apa_print(comp1_again_interaction)$full_result$Manipulation_HighRS_PrevsPost`. 
Those in the Control condition rated the investors post the exposure phase lower on all three attributes 
(Cooperation:`r papaja::apa_print(comp1_coop_interaction)$full_result$Control_HighRS_PrevsPost`, 
Forgiveness: `r papaja::apa_print(comp1_forg_interaction)$full_result$Control_HighRS_PrevsPost`, 
Play again: `r papaja::apa_print(comp1_again_interaction)$full_result$Control_HighRS_PrevsPost`).

For participants low on Rejection Sensitivity, those in the Manipulation condition rated the co-players in the "post" phase lower on all three attributes 
(Cooperation:`r papaja::apa_print(comp1_coop_interaction)$full_result$Manipulation_LowRS_PrevsPost`, 
Forgiveness: `r papaja::apa_print(comp1_forg_interaction)$full_result$Manipulation_LowRS_PrevsPost`, 
Play again: `r papaja::apa_print(comp1_again_interaction)$full_result$Manipulation_LowRS_PrevsPost`). 
Those in the Control condition did not differ in their ratings of the "pre" and "post" phase players.

In summary, we again see that those with low RS accurately perceive the co-player as similar on all attributes throughout the phases in the control condition. In contrast, the high RS group shows a negative bias towards the co-players after the "pre" phase in the control condition even though the player continues to use the same strategy. After exposure to the forgiving HMM, both groups  rates the "post" co-player worse than the "pre" even though they are the same.

@@

```{r}
saveRDS(summary(mod_returns_pct_RS), "data/mod_returns_pct_RS.RDS")
```

```{r}
contrasts(prepost_df$phase.f)
contrasts(prepost_df$condition.f)
contrasts(prepost_df$high_RS)

```
```{r}
RS_model <- lm( RS_score ~ ret_pct_na*inv_scaled + (1 + phase.f| playerId), prepost_df, REML= TRUE, method="KR")
```

@@

@@

<!-- EFFECT SIZES -->
```{r}
# Load the necessary packages

library(sjstats)
effectsize::eta_squared(mod_returns_pct_RS)
```

@@

Mixed-effects analysis on the percentage returns shows a significant main effect of Phase (Pre vs. Post RTG game), `r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f`, with higher percentage returns in the first RTG compared to the second. Importantly, we also find an interaction between Condition and Phase (RTG pre- vs. post-manipulation), `r papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f`. As shown in Figure \ref{fig:violinPlots}, post-hoc tests confirm a decrease in the percentage returned only in the manipulation condition, pre - post,
`r papaja::apa_print(prepost_bycond)$full_result$Manipulation_Pre_post`, but no change in the control condition.  We find no interaction between Phase, Condition and RS, suggesting that there was no difference between RS groups for this interaction.

```{r gamesPlot, include=T, echo=FALSE, fig.cap="Averages and standard errors of the trustee's return as a percentage of the multiplied investment received by Condition, Phase, and game round. The blue line shows the returns pre-manipulation and the green line post-manipulation. We note a different reaction to the pre-programmed one-off low investment between the two conditions: Whilst there is a dip in returns pre-manipulation for both conditions,  post manipulation we see higher returns in the manipulation condition compared to the dip in returns seen in the control condition in the right panel",fig.align="center", fig.width=6, fig.height = 4}

@@

# Plot
ggplot(avg_ret_df %>% filter(gameNum.f %in% c("pre","post")), aes(x=as.factor(roundNum), y=ret_pct_na, group=gameNum.f, color = gameNum.f, fill=gameNum.f)) +
  stat_summary(fun = mean, geom = "line") +
  stat_summary(fun.data = mean_se, geom = "ribbon", aes(ymin=..ymin.., ymax=..ymax..),
               alpha = 0.3, linetype = 0) +
  geom_vline(data = vline_data, aes(xintercept = xintercept, linetype = Defection_round), color = "black", linewidth = 0.6) +
  scale_linetype_manual(values = c("Pre-manipulation" = "dotted", "Post-manipulation" = "dashed"),name = "Defection round") +
  labs(x = "Round",
       y = "Percentage Return",
       color = "Trust Game Phase") +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.box = "vertical", # This will stack the legends vertically
        legend.key = element_blank(),
        legend.margin = margin(t = 0.2, b = 0.2, unit = "pt"),
        legend.box.margin = margin(t = 0, b = 0, unit = "pt"),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10)) +
  scale_color_manual(values = c("darkblue", "darkgreen"),
                     labels = c("Pre-manipulation", "Post-manipulation")) +
  scale_fill_manual(values = c("darkblue", "darkgreen"),
                    guide = "none") + # Hide the fill legend
  guides(color = guide_legend(override.aes = list(fill = NA)))+  # Remove fill for color legend keys
  guides(color = guide_legend(title = "Participant Returns Proportions", override.aes = list(fill = NA)),
         linetype = guide_legend(title = "Defection round")) +
  facet_wrap(~condition.f)

@@

There was also a significant main effect of Investment, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$inv_scaled)`,
such that higher investments were associated with higher percentage returns indicating positive reciprocity. An Investment by Condition interaction, 
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$condition_f_inv_scaled)`, reflected  that returns were more affected by investments in the control condition. We also find a three way interaction between Phase,  Investment and RS
`r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$gameNum_f_condition_f_inv_scaled)`, showing that the differentiated effect of the investment on the proportion returned by RS group is itself moderated by the Phase (pre- vs post manipulation). Finally, we find a four-way interaction between Condition, Phase, Investment and RS `r gsub("(\\d{1,3}),(\\d{3})", "\\1\\2", papaja::apa_print(mod_returns_pct_RS)$full_result$phase_f_condition_f_inv_scaled_high_RS)`.

```{r fourwayInteractionPlot, include=T, fig.cap ="Effect of investment on returns by condition, phase and RS group" }

library(ggplot2)

# 
# ggplot(prepost_df, aes(x = inv_scaled, y = ret_pct_na, color = phase.f)) +
#   # geom_point(alpha = 0.6) + # Adjust point transparency with alpha
#   geom_smooth(method = "lm", se = FALSE) + # Add linear regression lines without confidence intervals
#   labs(title = "Effect of inv_scaled on Outcome by Condition",
#        x = "inv_scaled", y = "Outcome") +
#   theme_minimal() +
#   facet_wrap(~high_RS)

# 'inv_scaled' is the continuous predictor, and 'outcome' is your dependent variable.
ggplot(prepost_df, aes(x = inv_scaled, y = ret_pct_na, color = phase.f)) +
  # geom_point(alpha = 0.6) + # Adjust point transparency with alpha
  geom_smooth(method = "lm", se = FALSE) + # Add linear regression lines without confidence intervals
  labs(x = "Investment Z-values", y = "Return proportion", color = "Experiment Phase") +
  theme_minimal() +
  facet_wrap(~condition.f*high_RS)+
  theme(legend.position = "bottom")

To interpret this interaction we plot the effect of the Investment on returns for each level of Condition, Phase and RS. Figure \ref{fig:fourwayInteractionPlot} shows that while the effect of investment on returns is similar across RS groups and across phases in the Control condition, the Manipulation condition shows a higher effect of investment in the post phase compared to pre in the low RS group, but a lower effect of investment compared to before for the high RS group. This suggests that high RS participants based their return less on the investment after being exposed to the forgiving AI compared to the low RS group.
